---
title: "Figures"
author: "Balazs Koncz"
date: "2024-04-22"
output: html_document
---

#Setup

```{r}
Packages <- c("readr", "tidyverse", "data.table", "magrittr", "fastmatch", "readxl")
#Packages <- c("magrittr", "dplyr", "data.table", "stringr", "ggplot2", "Rfast", "fastmatch", "forcats", "ggpubr", "tidyr", "ggseqlogo", "protr", "ComplexHeatmap", "readxl", "pbapply", "Hmisc", "seqinr")
# Packages <- c("magrittr", "dplyr", "data.table", "tidyr", , "Hmisc", , "tibble", "protr", "ComplexHeatmap", "stringr", "readr")
lapply(Packages, require, character.only = TRUE)
rm(Packages)
options(dplyr.summarise.inform = FALSE)

```

#I. All T cell assays (pos + neg)

```{r}
firstrow <- read_csv("D:/CloudStation/iedb/2023_04_11/tcell_full_v3.csv", n_max = 0) %>% names() #https://www.iedb.org/database_export_v3.php
firstrow = gsub("\\..*", "", firstrow)
firstrow = gsub(" ", "\\.", firstrow)
tcell_full = fread("D:/CloudStation/iedb/2023_04_11/tcell_full_v3.csv", skip = 1, check.names = T) #456,915
header = paste(firstrow, str_replace_all(string = colnames(tcell_full), pattern = "\\.[:digit:]+", replacement = ""), sep = ".")
colnames(tcell_full) = header

selcol = match(c("Epitope.Source.Organism", "Epitope.Species", "1st.immunogen.Name", "1st.immunogen.Reference.Name", "1st.immunogen.Source.Organism", "1st.immunogen.Species", "in.vitro.immunogen.Source.Organism", "in.vitro.immunogen.Species", "Assay.Antigen.Source.Organism", "Assay.Antigen.Species"), colnames(tcell_full))

tcell_filt_dengue = tcell_full %>% 
  filter_at(.vars = selcol, .vars_predicate = any_vars(grepl("dengue", ., ignore.case = T))) %>% #13,014
  filter(nchar(Epitope.Name) == 9) %>% #2087
  filter(grepl("HLA-A\\*|HLA-B\\*|HLA-C\\*", MHC.Restriction.Name)) %>%  #1630
  select(MHC.Restriction.Name, Epitope.Name) %>% 
  set_colnames(c("allele", "peptide")) %>% 
  mutate(organism = "Dengue")

tcell_filt_sars = tcell_full %>% 
  filter_at(.vars = selcol, .vars_predicate = any_vars(grepl("SARS-CoV2|Severe acute respiratory syndrome coronavirus 2", ., ignore.case = T))) %>% #20,125
  filter(nchar(Epitope.Name) == 9) %>% #5594
  filter(grepl("HLA-A\\*|HLA-B\\*|HLA-C\\*", MHC.Restriction.Name)) %>%  #4,632
  select(MHC.Restriction.Name, Epitope.Name) %>% 
  set_colnames(c("allele", "peptide")) %>% 
  mutate(organism = "SARS")

tcell_filt = rbind(tcell_filt_dengue, tcell_filt_sars)

rm(tcell_full, tcell_filt_dengue, tcell_filt_sars, firstrow, header, selcol)

pathogen_peptides = c("peptide", sort(unique(tcell_filt$peptide)))
writeLines(pathogen_peptides, "objects/pathogen_peptides.txt")
```

##Similarity calculation
###BLOSUM62 sequence similarity (full peptide, BL62 max, median)
####Functions

```{r}
create_db_proteome <- function(blast_folder, proteome_fasta_loc, peptide_length, db_name) {
  proteome = protr::readFASTA(proteome_fasta_loc)
  mers = lapply(proteome, FUN = function(z) substring(z, 1:(nchar(z)-(peptide_length - 1)), peptide_length:nchar(z)))
  mers = unique(unlist(mers, use.names = F))
  strings = rbind(paste0(">", mers), mers)
  strings = as.vector(strings)
  writeLines(strings, paste0(blast_folder, db_name, ".fasta"))
  message("Makeblastdb application is curently produces BLAST databases from your proteome...")
  system2(command = paste0(blast_folder, "makeblastdb.exe"), args = c(paste0("-in ", blast_folder, db_name, ".fasta"), "-dbtype prot", "-title epitopes"))
  message(paste0("Database is created in ", blast_folder))
}

blastp <- function(blast_folder, query_peptides, db_loc, out_loc, threads = 1, keep_query_fasta_peptides = "NO") {
  strings = rbind(paste0(">", query_peptides), query_peptides)
  strings = as.vector(strings)
  writeLines(strings, paste0(blast_folder, "query_peptides.fasta"))
  #message("blastp is already running...")
  system2(command = paste0(blast_folder, "blastp.exe"), 
          args = c(paste0("-query ", paste0(blast_folder, "query_peptides.fasta")), 
                   paste0("-db ", db_loc), 
                   paste0("-out ", out_loc),
                   paste0("-num_threads ", threads),
                   "-outfmt 6",
                   "-task blastp-short",
                   "-ungapped",
                   "-comp_based_stats F",
                   "-evalue 10000000",
                   "-max_target_seqs 100"))
  #message(paste0("Sequences were compared. The location of the output file: ", out_loc))
  if(keep_query_fasta_peptides == "NO") unlink(paste0(blast_folder, "query_peptides.fasta"))
}

calcSeqSim = function(blastpout_loc) {
  aLign = function(seq_1, seq_2, mtx = protr::AABLOSUM62, aas = colnames(protr::AABLOSUM62)) {
    sum(mtx[cbind(fmatch(unlist(strsplit(seq_1, "")), aas), fmatch(unlist(strsplit(seq_2, "")), aas))])
  }
  #library(pbapply)
  library(fastmatch)
  library(future.apply)
  blastpout = read.table(blastpout_loc, header = F, stringsAsFactors = F)
  mers = unique(blastpout$V1)
  future_sapply(mers, function(x) {
    max(sapply(blastpout$V2[blastpout$V1 == x], function(y) {
      aLign(x, y)/(sqrt((aLign(x, x))*(aLign(y, y))))
    }), na.rm = T)
  })
}

calcSeqSimMED = function(blastpout_loc) {
  aLign = function(seq_1, seq_2, mtx = protr::AABLOSUM62, aas = colnames(protr::AABLOSUM62)) {
    sum(mtx[cbind(fmatch(unlist(strsplit(seq_1, "")), aas), fmatch(unlist(strsplit(seq_2, "")), aas))])
  }
  library(fastmatch)
  library(future.apply)
  blastpout = read.table(blastpout_loc, header = F, stringsAsFactors = F)
  mers = unique(blastpout$V1)
  future_sapply(mers, function(x) {
    median(sapply(blastpout$V2[blastpout$V1 == x], function(y) {
      aLign(x, y)/(sqrt((aLign(x, x))*(aLign(y, y))))
    }), na.rm = T)
  })
}

```

####Calculation

```{r}
create_db_proteome(blast_folder = "C:/NCBI/ncbi-blast-2.12.0+/bin/", proteome_fasta_loc = "objects/uniprot-compressed_true_download_true_format_fasta_query__28proteome-2023.04.11-08.26.57.89.fasta", peptide_length = 9, db_name = "human_proteome_20230411_9mers")

peps = unique(tcell_filt$peptide)
blastp(blast_folder = "C:/NCBI/ncbi-blast-2.12.0+/bin/", query_peptides = peps, db_loc = "C:/NCBI/ncbi-blast-2.12.0+/bin/human_proteome_20230411_9mers.fasta", out_loc = "objects/blastpout_ss_iedb_dengue_sars_peps2261", threads = 6)

ss_scores = calcSeqSim(blastpout_loc = "objects/blastpout_ss_iedb_dengue_sars_peps2261")
save(ss_scores, file = "objects/sequence_similarity_values_peps2261")
ss_scores = calcSeqSimMED(blastpout_loc = "objects/blastpout_ss_iedb_dengue_sars_peps2261")
save(ss_scores, file = "objects/sequence_similarityMED_values_peps2261")

```

###Dissimilarity


```{r}
v = unique(tcell_filt$peptide)
v = v[sapply(v, protr::protcheck)]
names(v) <- 1:length(v) %>% as.character()

sdt <- v %>%
  data.table::as.data.table() %>%
  .[, nmer_id := names(v)]

AA <- Biostrings::AAStringSet(v, use.names = TRUE)
Biostrings::writeXStringSet(AA, file = "C:/NCBI/ncbi-blast-2.12.0+/bin/dissimilarity_fasta_iedb_dengue_sars_peps2261.fa", format = "fasta")


# blast call with the parameters from Luksza model, E value, matrix, and gap costs, https://github.com/leeprichman/Richman_2019_Cell_Systems/blob/master/Figure%203/Fig3CDE.R
#blastp -query dissimilarity_fasta_iedb_dengue_sars_peps2261.fa -db human_proteome_9mers.fasta -out iedb_dengue_sars_peps2261_dissim.csv -num_threads 7 -matrix BLOSUM62 -gapopen 11 -gapextend 1 -evalue 100000000 -outfmt "10 qseqid sseqid qseq qstart qend sseq sstart send length mismatch pident evalue bitscore"

blastdt <- "C:/NCBI/ncbi-blast-2.12.0+/bin/iedb_dengue_sars_peps2261_dissim.csv" %>% fread

blastdt %>% data.table::setnames(names(.),
                                 c("nmer_id",
                                   "self_anno",
                                   "nmer",
                                   "q_start",
                                   "q_stop",
                                   "WT.peptide",
                                   "s_start",
                                   "s_end",
                                   "overlap_length",
                                   "mismatch_length",
                                   "pident",
                                   "evalue",
                                   "bitscore"))
# shrink table for memory constraints
blastdt <- blastdt[, .SD %>% unique, .SDcols = c("nmer_id", "self_anno", "WT.peptide", "nmer")]

# remove gapped alignments, necessary given permissive blast params
blastdt <- blastdt[, WT.peptide := WT.peptide %>% stringr::str_replace_all(pattern = "-|\\*|U", replacement = "")] %>%
  .[, nmer := nmer %>% stringr::str_replace_all(pattern = "-|\\*|U", replacement = "")] %>%
  .[!is.na(WT.peptide) & !is.na(nmer)] %>%
  .[nchar(nmer) == nchar(WT.peptide)]

# get table of alignment lengths per nmer
blastdt[, align_l := nchar(WT.peptide)]


# this is memory intense, lets split it up and stitch it back
suppressWarnings(
  blastdt <- blastdt %>% split(1:(nrow(blastdt) / 100))
)

blastdt <- lapply(blastdt %>% seq_along(), function(i) {
  fn <- paste("objects/blastdt_obs/blastdt_", i, ".txt", sep = "")

  blastdt[[i]] %>% data.table::fwrite(fn, sep = "\t")

  return(fn)
}) %>% unlist()



make_sw_alignment <- function(col1,
                              col2,
                              gap_open = -11,
                              gap_extend = -1) {
  al <- Biostrings::pairwiseAlignment(col1, col2,
                                      substitutionMatrix = "BLOSUM62",
                                      gapOpening = gap_open,
                                      gapExtension = gap_extend,
                                      type = "local",
                                      scoreOnly = TRUE
  )

  if (length(al) == 0) al <- as.numeric(NA)

  return(al)
}

pblapply(blastdt %>% seq_along(), function(i) {

  print(paste("Alignment subset", i, "of", length(blastdt)))

  b <- blastdt[i] %>% data.table::fread()

  b[, SW := make_sw_alignment(nmer, WT.peptide)]

  b %>% data.table::fwrite(blastdt[i], sep = "\t")

  return(NULL)
})

blastdt <- lapply(blastdt, function(f) {
  dt <- data.table::fread(f)

  file.remove(f)

  return(dt)
}) %>% data.table::rbindlist(use.names = TRUE, fill = TRUE)

#message("Running partition function...")


modeleR <- function(als, a = aval, k = kval) {
  be <- -k * (a - als)
  sumexp <- sum(exp(be))
  Zk <- 1 + sumexp
  R <- sumexp / Zk
  R <- 1 - R
  return(R)
}

kval = 4.86936
aval = 32 #I modified this value from 34 to 32 (according to the Luksza article) https://github.com/andrewrech/antigen.garnish/blob/main/R/antigen.garnish_predict.R

blastdt[, dissimilarity := SW %>% modeleR(), by = "nmer_id"]

blastdt <- blastdt[, .SD %>% unique(), .SDcols = c("nmer_id", "dissimilarity")]

sdt[, nmer_id := as.character(nmer_id)]
blastdt[, nmer_id := as.character(nmer_id)]

sdt <- merge(sdt, blastdt, by = "nmer_id")

sdt %>% data.table::setnames(".", "nmer")

sdt = sdt[, .SD %>% unique(), .SDcols = c("nmer", "dissimilarity")]
save(sdt, file = "objects/dissimilarity_values")

rm(AA, blastdt, sdt, aval, kval, v, make_sw_alignment, modeleR)
```

###Mayer
Nb. of mismatched amino acids

```{r}
blastpout = fread("objects/blastpout_ss_iedb_dengue_sars_peps2261")
blastpout2 = blastpout %>% group_by(qseqid) %>% slice(which.min(strdist))
save(aadifs, file = "objects/aadifs_mayer")
rm(blastpout, mers, aadifs)
```

###TCEMfreq

```{r}
proteome = protr::readFASTA("objects/uniprot-compressed_true_download_true_format_fasta_query__28proteome-2023.04.11-08.26.57.89.fasta")
mers9 = lapply(proteome, FUN = function(z) substring(z, 1:(nchar(z)-(9 - 1)), 9:nchar(z)))
mers9 = unlist(mers9, use.names = F)
mers9 = mers9[nchar(mers9) == 9]
tcems = substr(mers9,4,8)
tcems = tcems[!grepl("U", tcems)]
tcemfreq = Table(tcems)
save(tcemfreq, file = "objects/tcemfreq")
rm(proteome, mers9, tcemfreq, tcems)
```

##Add to dataset

```{r}
tcell_filt = unique(tcell_filt)

load("objects/sequence_similarity_values_peps2261")
tcell_filt$bl62ssMax = ss_scores[fmatch(tcell_filt$peptide, names(ss_scores))]
rm(ss_scores)

load("objects/sequence_similarityMED_values_peps2261")
tcell_filt$bl62ssMed = ss_scores[fmatch(tcell_filt$peptide, names(ss_scores))]
rm(ss_scores)

load("objects/dissimilarity_values")
tcell_filt$dissim = sdt$dissimilarity[fmatch(tcell_filt$peptide, sdt$nmer)]
rm(sdt)

load("objects/aadifs_mayer")
tcell_filt$aadif = aadifs[fmatch(tcell_filt$peptide,names(aadifs))]
rm(aadifs)

tcell_filt$exactmatch = ifelse(tcell_filt$aadif == 0, 1, 0)

load("objects/tcemfreq")
tcell_filt$tcemfreq = tcemfreq[fmatch(substr(tcell_filt$peptide,4,8), names(tcemfreq))]
tcell_filt$tcemfreq[is.na(tcell_filt$tcemfreq)] = 0
rm(tcemfreq)

save(tcell_filt, file = "objects/tcell_filt")
```

#II. Neoepitopes

##Dataset

```{r}
tesla_data = read_excel("mmc4.xlsx") #https://doi.org/10.1016/j.immuni.2023.09.002
tesla_data %<>% select(MHC, ALT_EPI_SEQ) %>% filter(nchar(ALT_EPI_SEQ) == 9) %>% set_colnames(c("allele", "peptide"))
neopeptides_tesla = c("peptide", sort(unique(tesla_data$peptide)))
writeLines(neopeptides_tesla, "objects/neopeptides_tesla.txt")
```

##Similarity calculation
###BLOSUM62 sequence similarity (full peptide, BL62 max, median)

```{r}
peps = unique(tesla_data$peptide)

blastp(blast_folder = "C:/NCBI/ncbi-blast-2.12.0+/bin/", query_peptides = peps, db_loc = "C:/NCBI/ncbi-blast-2.12.0+/bin/human_proteome_20230411_9mers.fasta", out_loc = "objects/blastpout_ss_tesla_neoeps_301", threads = 6)
ss_scores = calcSeqSim(blastpout_loc = "objects/blastpout_ss_tesla_neoeps_301")
save(ss_scores, file = "objects/sequence_similarity_values_neoeps301")
ss_scores = calcSeqSimMED(blastpout_loc = "objects/blastpout_ss_tesla_neoeps_301")
save(ss_scores, file = "objects/sequence_similarityMED_values_neoeps301")

```

###Dissimilarity

```{r}
v = unique(tesla_data$peptide)
v = v[sapply(v, protr::protcheck)]
names(v) <- 1:length(v) %>% as.character()

sdt <- v %>%
  data.table::as.data.table() %>%
  .[, nmer_id := names(v)]

AA <- Biostrings::AAStringSet(v, use.names = TRUE)
Biostrings::writeXStringSet(AA, file = "C:/NCBI/ncbi-blast-2.12.0+/bin/dissimilarity_fasta_tesla_neoeps301.fa", format = "fasta")


# blast call with the parameters from Luksza model, E value, matrix, and gap costs, https://github.com/leeprichman/Richman_2019_Cell_Systems/blob/master/Figure%203/Fig3CDE.R
#blastp -query dissimilarity_fasta_tesla_neoeps301.fa -db human_proteome_9mers.fasta -out tesla_neoeps301_dissim.csv -num_threads 7 -matrix BLOSUM62 -gapopen 11 -gapextend 1 -evalue 100000000 -outfmt "10 qseqid sseqid qseq qstart qend sseq sstart send length mismatch pident evalue bitscore"

blastdt <- "C:/NCBI/ncbi-blast-2.12.0+/bin/tesla_neoeps301_dissim.csv" %>% fread

blastdt %>% data.table::setnames(names(.),
                                 c("nmer_id",
                                   "self_anno",
                                   "nmer",
                                   "q_start",
                                   "q_stop",
                                   "WT.peptide",
                                   "s_start",
                                   "s_end",
                                   "overlap_length",
                                   "mismatch_length",
                                   "pident",
                                   "evalue",
                                   "bitscore"))
# shrink table for memory constraints
blastdt <- blastdt[, .SD %>% unique, .SDcols = c("nmer_id", "self_anno", "WT.peptide", "nmer")]

# remove gapped alignments, necessary given permissive blast params
blastdt <- blastdt[, WT.peptide := WT.peptide %>% stringr::str_replace_all(pattern = "-|\\*|U", replacement = "")] %>%
  .[, nmer := nmer %>% stringr::str_replace_all(pattern = "-|\\*|U", replacement = "")] %>%
  .[!is.na(WT.peptide) & !is.na(nmer)] %>%
  .[nchar(nmer) == nchar(WT.peptide)]

# get table of alignment lengths per nmer
blastdt[, align_l := nchar(WT.peptide)]


# this is memory intense, lets split it up and stitch it back
suppressWarnings(
  blastdt <- blastdt %>% split(1:(nrow(blastdt) / 100))
)

blastdt <- lapply(blastdt %>% seq_along(), function(i) {
  fn <- paste("objects/blastdt_obs/blastdt_", i, ".txt", sep = "")

  blastdt[[i]] %>% data.table::fwrite(fn, sep = "\t")

  return(fn)
}) %>% unlist()



make_sw_alignment <- function(col1,
                              col2,
                              gap_open = -11,
                              gap_extend = -1) {
  al <- Biostrings::pairwiseAlignment(col1, col2,
                                      substitutionMatrix = "BLOSUM62",
                                      gapOpening = gap_open,
                                      gapExtension = gap_extend,
                                      type = "local",
                                      scoreOnly = TRUE
  )

  if (length(al) == 0) al <- as.numeric(NA)

  return(al)
}

pblapply(blastdt %>% seq_along(), function(i) {

  print(paste("Alignment subset", i, "of", length(blastdt)))

  b <- blastdt[i] %>% data.table::fread()

  b[, SW := make_sw_alignment(nmer, WT.peptide)]

  b %>% data.table::fwrite(blastdt[i], sep = "\t")

  return(NULL)
})

blastdt <- lapply(blastdt, function(f) {
  dt <- data.table::fread(f)

  file.remove(f)

  return(dt)
}) %>% data.table::rbindlist(use.names = TRUE, fill = TRUE)

#message("Running partition function...")


modeleR <- function(als, a = aval, k = kval) {
  be <- -k * (a - als)
  sumexp <- sum(exp(be))
  Zk <- 1 + sumexp
  R <- sumexp / Zk
  R <- 1 - R
  return(R)
}

kval = 4.86936
aval = 32

blastdt[, dissimilarity := SW %>% modeleR(), by = "nmer_id"]

blastdt <- blastdt[, .SD %>% unique(), .SDcols = c("nmer_id", "dissimilarity")]

sdt[, nmer_id := as.character(nmer_id)]
blastdt[, nmer_id := as.character(nmer_id)]

sdt <- merge(sdt, blastdt, by = "nmer_id")

sdt %>% data.table::setnames(".", "nmer")

sdt = sdt[, .SD %>% unique(), .SDcols = c("nmer", "dissimilarity")]
save(sdt, file = "objects/dissimilarity_values_neoeps")

rm(AA, blastdt, sdt, aval, kval, v, make_sw_alignment, modeleR)
```

###Mayer
Nb. of mismatched amino acids

```{r}
blastpout = fread("objects/blastpout_ss_tesla_neoeps_301")
mers = unique(blastpout$V1)
aadifs = pbsapply(mers, function(x) {
  tempdf = blastpout[blastpout$V1 == x,]
  min(apply(tempdf, 1, function(z) sum(Reduce("!=", strsplit(c(z[1], z[2]), split = "")))))
})
save(aadifs, file = "objects/aadifs_mayer_neoeps")


# library(stringdist)
# colnames(blastpout) = c("qseqid", "sseqid", "pident", "length", "mismatch", "gapopen", "qstart", "qend", "sstart", "send", "evalue", "bitscore")
# blastpout$strdist = apply(blastpout, 1, function(x) {
#   stringdist(a = x[1], b = x[2], method = "hamming")
# })
# blastpout2 = blastpout %>% group_by(qseqid) %>% slice(which.min(strdist)) 
# 
# strdist = blastpout2$strdist
# names(strdist) = blastpout2$qseqid
# 
# View(cbind(aadifs, strdist[match(names(aadifs), names(strdist))]))

rm(blastpout, mers, aadifs)
```

##Add to dataset

```{r}
tesla_data = unique(tesla_data)

load("objects/sequence_similarity_values_neoeps301")
tesla_data$bl62ssMax = ss_scores[fmatch(tesla_data$peptide, names(ss_scores))]
rm(ss_scores)

load("objects/sequence_similarityMED_values_neoeps301")
tesla_data$bl62ssMed = ss_scores[fmatch(tesla_data$peptide, names(ss_scores))]
rm(ss_scores)

load("objects/dissimilarity_values_neoeps")
tesla_data$dissim = sdt$dissimilarity[fmatch(tesla_data$peptide, sdt$nmer)]
rm(sdt)

load("objects/aadifs_mayer_neoeps")
tesla_data$aadif = aadifs[fmatch(tesla_data$peptide,names(aadifs))]
rm(aadifs)

tesla_data$exactmatch = ifelse(tesla_data$aadif == 0, 1, 0)

load("objects/tcemfreq")
tesla_data$tcemfreq = tcemfreq[fmatch(substr(tesla_data$peptide,4,8), names(tcemfreq))]
tesla_data$tcemfreq[is.na(tesla_data$tcemfreq)] = 0
rm(tcemfreq)

save(tesla_data, file = "objects/tesla_data")
```

#III. FIGURES

##Fig1 Correlation plot
###C1 Pathogen

```{r}
load("objects/tcell_filt")
tcell_filt %<>% select(peptide, bl62ssMax, bl62ssMed, dissim, aadif, tcemfreq) %>% unique()
colnames(tcell_filt) = c("peptide", "BL62 similarity\nscore (max)", "BL62 similarity\nscore (med)", "Dissimilarity", "Hamming distance", "TCEM frequency")

cordf <- expand.grid(Var1 = colnames(tcell_filt)[2:6], Var2 = colnames(tcell_filt)[2:6], stringsAsFactors = F)
cordf = cordf[c(1,6:7,11:13,16:19,21:25),]
#cordf = cordf[c(1:6,8:12,15:18,22:24,29,30,36),]
cordf$sp_rho = unname(apply(cordf, 1, function(x) cor.test(tcell_filt %>% pull(x[1]), tcell_filt %>% pull(x[2]), type = "spearman")$estimate))
cordf$sp_pvalue = unname(apply(cordf, 1, function(x) cor.test(tcell_filt %>% pull(x[1]), tcell_filt %>% pull(x[2]), type = "spearman")$p.value))

#ALMOST GOOD
cordf$label = apply(cordf, 1, function(x) {
  if(as.numeric(x[3]) == 1) {
    out = ""
  } else if(as.numeric(x[4]) < 1e-3) {
    out = paste0("paste(\"",
           round(as.numeric(x[3]),2),
           "\n",
           "(",
           strsplit(format(as.numeric(x[4]),digits = 2, scientific = T), "e")[[1]][1],
           " x \",10 ^ ", 
           as.numeric(strsplit(format(as.numeric(x[4]), digits = 2, scientific = T), "e")[[1]][2]),
           ",\")\")")
  } else {
    out = paste0("paste(\" ", 
           round(as.numeric(x[3]),2), 
           "\n",
           "(", 
           round(as.numeric(x[4]),3), 
           ")", 
           "\")")
  }
  out
})

cordf$sp_rho[cordf$Var1 == cordf$Var2] = NA

cordf$Var1 = factor(cordf$Var1, levels = c("BL62 similarity\nscore (max)", "BL62 similarity\nscore (med)", "Dissimilarity", "Hamming distance", "TCEM frequency"))
cordf$Var2 = factor(cordf$Var2, levels = c("BL62 similarity\nscore (max)", "BL62 similarity\nscore (med)", "Dissimilarity", "Hamming distance", "TCEM frequency"))

fig_cor_path = ggplot(cordf, aes(Var2, Var1, fill = sp_rho)) +
  geom_tile(color = "black") +
  scale_fill_gradient2(low = "#3C5488B2", high = "#DC0000B2", mid = "white", midpoint = 0, limit = c(-1,1), space = "Lab", name = "Spearman Correlation", na.value = "white") +
  scale_x_discrete(position = "top") +
  scale_y_discrete(position = "left") +
  theme_minimal() + 
  coord_fixed() +
  geom_text(aes(Var2, Var1, label = label), color = "black", size = 3, parse = T) +
  labs(title = "Pathogen peptides") +
  theme(title = element_text(size = 14),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 12, angle = 90, hjust = 0),
        axis.text.y = element_text(size = 12, hjust = 1),
        panel.grid.major = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.ticks = element_blank(),
        legend.position.inside = "none")
fig_cor_path

```

###C2 Neoepitopes

```{r}
load("objects/tesla_data")
tesla_data %<>% select(peptide, bl62ssMax, bl62ssMed, dissim, aadif, tcemfreq) %>% unique()
colnames(tesla_data) = c("peptide", "BL62 similarity\nscore (max)", "BL62 similarity\nscore (med)", "Dissimilarity", "Hamming distance", "TCEM frequency")


cordf_tesla <- expand.grid(Var1 = colnames(tesla_data)[2:6], Var2 = colnames(tesla_data)[2:6], stringsAsFactors = F)
cordf_tesla = cordf_tesla[c(1,6:7,11:13,16:19,21:25),]
cordf_tesla$sp_rho = unname(apply(cordf_tesla, 1, function(x) cor.test(tesla_data %>% pull(x[1]), tesla_data %>% pull(x[2]), type = "spearman")$estimate))
cordf_tesla$sp_pvalue = unname(apply(cordf_tesla, 1, function(x) cor.test(tesla_data %>% pull(x[1]), tesla_data %>% pull(x[2]), type = "spearman")$p.value))

cordf_tesla$label = apply(cordf_tesla, 1, function(x) {
  if(as.numeric(x[3]) == 1) {
    out = ""
  } else if(as.numeric(x[4]) < 1e-3) {
    paste0("paste(\"",
           round(as.numeric(x[3]),2),
           "\n",
           "(",
           strsplit(format(as.numeric(x[4]),digits = 2, scientific = T), "e")[[1]][1],
           " x \",10 ^ ", 
           as.numeric(strsplit(format(as.numeric(x[4]), digits = 2, scientific = T), "e")[[1]][2]),
           ",\")\")")
  } else {
    paste0("paste(\" ", 
           round(as.numeric(x[3]),2), 
           "\n",
           "(", 
           round(as.numeric(x[4]),3), 
           ")", 
           "\")")
  }
})

cordf_tesla$sp_rho[cordf_tesla$Var1 == cordf_tesla$Var2] = NA

cordf_tesla$Var1 = factor(cordf_tesla$Var1, levels = c("BL62 similarity\nscore (max)", "BL62 similarity\nscore (med)", "Dissimilarity", "Hamming distance", "TCEM frequency"))
cordf_tesla$Var2 = factor(cordf_tesla$Var2, levels = c("BL62 similarity\nscore (max)", "BL62 similarity\nscore (med)", "Dissimilarity", "Hamming distance", "TCEM frequency"))
fig_cor_neoep = ggplot(cordf_tesla, aes(Var2, Var1, fill = sp_rho)) +
  geom_tile(color = "black") +
  scale_fill_gradient2(low = "#3C5488B2", high = "#DC0000B2", mid = "white", midpoint = 0, limit = c(-1,1), space = "Lab", name = "Spearman Correlation", na.value = "white") +
  scale_x_discrete(position = "top") +
  scale_y_discrete(position = "left") +
  theme_minimal() + 
  coord_fixed() +
  geom_text(aes(Var2, Var1, label = label), color = "black", size = 3, parse = T) +
  labs(title = "Neopeptides") +
  theme(title = element_text(size = 14),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 12, angle = 90, hjust = 0),
        axis.text.y = element_text(size = 12, hjust = 1),
        panel.grid.major = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.ticks = element_blank(), 
        legend.title = element_text(size = 12, angle = 90, vjust = 1, hjust = .5),
        legend.position.inside = c(0.9, 0.16)) +
        #legend.justification = c(1, 0)),
  guides(fill = guide_colorbar(title = "Spearman's rho", title.position = "left", direction = "vertical", barwidth = 1, barheight = 7))
fig_cor_neoep

ggarrange(plotlist = list(fig_cor_path, fig_cor_neoep), ncol = 2, nrow = 1, common.legend = T, legend = "right")

ggsave(filename = "figures/fig1c_v5.jpg", width = 34, height = 19, units = "cm", dpi = "retina")
```

##Fig1D Scores per allele

```{r}
load("objects/tcell_filt")
load("objects/tesla_data")

tesla_data$allele = paste0("HLA-", tesla_data$allele)
tcell_filt %<>% select(allele, peptide, bl62ssMax, dissim)
tesla_data %<>% select(allele, peptide, bl62ssMax, dissim)

data = rbind(tcell_filt, tesla_data)
rm(tcell_filt, tesla_data)

data %<>% mutate(locus = substr(allele, 5, 5))
selalleles = data %>% add_count(allele, name = "allelefreq") %>% filter(allelefreq > 50) %>% pull(allele) %>% unique()

```

###BL62ssmax

```{r}
tempdf_bl62max = data %>% 
  filter(allele %in% selalleles) %>% 
  mutate(allele = fct_reorder(allele, bl62ssMax, median))
kruskal.test(bl62ssMax~allele, tempdf_bl62max)$p.value

ggplot(tempdf_bl62max, aes(x = allele, y = bl62ssMax, fill = locus)) + 
  geom_boxplot(outlier.shape = NA, alpha=0.5) +
  geom_jitter(mapping = aes(color = locus), size=0.4, alpha=0.9) +
  annotate(geom = "text", x = 4.5, y = 1, label = "paste(italic(P), \" = 7.0 x \", 10 ^ -26)", parse = TRUE, color = "black", size = 4) +
  scale_fill_manual(values = alpha(c("#DC0000B2","#3C5488B2"), .3)) +
  scale_color_manual(values = c("#DC0000B2","#3C5488B2")) +
  ylab("BLOSUM62 similarity score (max)") +
  theme_bw() +
  theme(
    legend.position = "none",
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 14),
    axis.text.x = element_text(size = 12, color = "black", angle = 90, vjust = .5),
    axis.text.y = element_text(size = 12, color = "black"),
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(colour = "black"),
    plot.margin = margin(7, 0, 6.5, 2, "cm")) -> fig_bl62ssmax



```

###Dissimilarity

```{r}
tempdf_dis = data %>% 
  filter(allele %in% selalleles) %>% 
  mutate(allele = fct_reorder(allele, dissim, median))
kruskal.test(dissim~allele, tempdf_dis)$p.value
tempdf_dis %<>% mutate(dissimM = dissim + min(tempdf_dis$dissim[tempdf_dis$dissim != 0]))

ggplot(tempdf_dis, aes(x = allele, y = dissimM, fill = locus)) + 
  geom_boxplot(outlier.shape = NA, alpha=0.5) +
  geom_jitter(mapping = aes(color = locus), size = 0.4, alpha = 0.9) +
  annotate(geom = "text", x = 4.5, y = 10, label = "paste(italic(P), \" = 4.2 x \", 10 ^ -5)", parse = TRUE, color = "black", size = 4) +
  scale_fill_manual(values = alpha(c("#DC0000B2","#3C5488B2"), .3)) +
  scale_color_manual(values = c("#DC0000B2","#3C5488B2")) +
  scale_y_continuous(trans = "log10", breaks = c(1e-14,1e-12,1e-10,1e-8,1e-6,1e-4,0.01,1), labels = c(10^-14, 10^-12, 10^-10, 10^-8, 10^-6, 10^-4, 0.01, 1)) +
  ylab("Dissimilarity") +
  theme_bw() +
  theme(
    legend.position = "none",
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 14),
    axis.text.x = element_text(size = 12, color = "black", angle = 90, vjust = .5),
    axis.text.y = element_text(size = 12, color = "black"),
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(colour = "black"),
    plot.margin = margin(7, 0, 6.5, 2, "cm")) -> fig_dissim

ggarrange(plotlist = list(fig_bl62ssmax, fig_dissim), ncol = 1, nrow = 2)

ggsave(filename = "figures/Fig1d_v2.jpg", width = 28, height = 28, units = "cm", dpi = "retina")

```

